---
title: "R Notebook-CLASS-1"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*.

```         
31*78
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

# Ex-1

```{r}
a <- 5*3
a
```

```{r}
help(log)
```

```{r}
a<-1
b<- (-6)
c<-1
s1<- (-b+sqrt(b^2 - 4*a*c))/(2*a)
s2<- (-b-sqrt(b^2 - 4*a*c))/(2*a)
print(s1)
print(s2)
```

# Vectros

```{r}
vector1<- c(1,2,3)
vector1
class(vector1)
vect2<- c('i','n','d','i','a')
vect2
class(vect2)
vect3<-c(1,2,'i','p')
vect3
class(vect3)
vect4<-c("italy"=10,"canada"=20,"usa"=30 )
vect4
class(vect4)
seq(1,10)
1:5
names(vect4)
vect4["italy"]
vect4[2]
vect4[1:3]
vect4[c(1,3)]

```

```{r}
a<- c(1,"canada",2)
class(a)
vect3<-c(1,2,'i','v')
class(vect3)
```

```{r}
x<- 1:10
x
y<- as.character(x)
y
as.numeric(y)
```

# matrix

```{r}
mat<- matrix(1:12,nrow = 4,ncol = 3)
mat
mat[2,2]
mat[2,]
mat[,3]
mat[,2:3]
mat[1,2:3]


```

```{r}
as.data.frame(mat)
```

# list

```{r}
list1=list(1,2,3)
list1
mylist1=list(name="prabal",degree="msc",course="dsai")
mylist1
out=t.test(1:10,c(7:20))
out
```

```{r}
a<-c(1,2,3,4)
b<-c("d","e","f","g")
c<-c("hi","bp","np","gh")
df<-data.frame(a,b,c)
df
df[1:3,1:2]
df[1,2]
df[1,]
df[,3]
```

```{r}
new_c1<- c("kol","hol","jol","pol")
df$city<- new_c1
df
subset(df,a>2)
```

# Import Library

```{r}
library(dslabs)
data(murders)
class(murders)
```

```{r}
str(murders)
```

```{r}
head(murders)
```

```{r}
names(murders)
```

```{r}
murders$population
```

```{r}
plot(1:10,(1:10)^4)
```

```{r}
plot(1:10,(1:10)^4,xlab = 'val1',ylab = 'val2',pch=19,type = 'b',col='red',main = "Plot1",cex=3,lty=3)

```

```{r}
library(RColorBrewer)
barplot(1:4,col=brewer.pal(4,"Set1"))
```

```{r}
barplot(1:4,col=c("red","green"))


```

# Scatter plot

```{r}
x<- murders$population/10^6
y<- murders$total
plot(x,y,xlim = c(0,40), ylim =c(0,1500),pch=19,xlab = 'population',ylab = 'total',col='red',main = "Murders")
legend("bottomright",legend = "data points",pch=19,col = "red")
abline(a=0,b=30,lty=2,lwd=2.5,col=("blue"))
```

# BOX PLOT

```{r}
murders$rate<- with(murders,total/population * 100000)
head(murders)

boxplot(rate~region, data = murders)
```

# Histogram

```{r}
x<- with(murders,total/population * 100000)
hist(x)
murders$state[which.max(x)]

```

```{r}
x= runif(100,0,1)
par(mflow=c(1,2))
hist(x)
hist(x, breaks=2)
hist(x, breaks=100)

```

# Density plot

```{r}

x= runif(100,0,1)

hist(x, freq = FALSE)
# hist(x, breaks=2)
# hist(x, breaks=100)
f=density(x)
lines(f,col="red",lty=2,lwd=2.5)
```

# QQ PLOT

```{r}

x= runif(100,0,1)
qqnorm(x)
```

# pie charts

```{r}

pie(summary(murders$region))
```

# Pair Plot

```{r}

x<-murders[,4:6]
pairs(x,col= as.numeric(murders$region),pch=as.numeric(murders$region))
```
IF ELSE AND LOOPS 
===
```{r}
x<-10
if(x<20){
  print("hi i am in france")
}
```

```{r}

x<-10
if(x>20){
  print("hi i am in france")
}else{
  print(" Hi I am in USA")
}
```
```{r}

x<-9
if(x>20){
  print("hi i am in france")
}else if(x <10){
  print(" I am in Germany")
  }else{
  print(" Hi I am in USA")
}
```
```{r}
runif(1,0,10)
```

```{r}

x<-9
y<- if(x>20){
  "hi i am in france"
}else if(x <10){
  " I am in Germany"
  }else{
  " Hi I am in USA"
  }
y
```
```{r}

for(i in 1:10){
  print(i)
}
```

```{r}
count<-10

while(count>1){
  print(count)
  count<- count-1
}
```

