---
title: "PRABAL GHOSH R EXAM-1"
output:
  pdf_document: default
  html_document:
    df_print: paged
---



#Part I – Explore the data

```{r}

library(ggplot2)
```

```{r}
library("dplyr")
```



```{r}
data_small <- read.csv('C:\\Users\\praba\\Desktop\\uca1\\M1\\R programming\\exam1\\people_habits_dataset.csv')

```


```{r}
data_small
```



##1.What are the dimensions of the data?

```{r}
dim(data_small)
```

##2.What are the types of columns? What is the difference between a string and a factor? Show descriptive statistics of the data.

If x is a string it can take any value. If x is a factor it can only take a values from a list of all levels. That makes these variables more memory effecient as well.

```{r}
str(data_small)
```



```{r}
summary(data_small)
```


##3.Is there any NA?


```{r}
which(is.na(data_small))
print ("Row and Col positions of NA values") 
which(is.na(data_small), arr.ind=TRUE)
```


#Part II – Data transformation



##4.Create a new variable called age from the variable dob to obtain the age of individuals based on their date of birth. (Hint: you can use the function year() from lubridate package to get the year of the date of birth).



```{r}
 
# install.packages("lubridate")

```
```{r}
library(lubridate)
```



```{r}
choices = lubridate::year(Sys.Date())
data_small_1<- data_small %>% mutate(data_small, age= (choices-year(data_small$dob) ))
data_small_1

```


##5.Calculate the body mass index (BMI) for each of our patients using the formula BMI=Weight/(Height)2, where the weight is measured in kilograms and the height is in meters


```{r}
bmi <- function(weight, height) {
height/100 -> height
height^2 -> height
return(weight/height)
}
bmi_result <- bmi(data_small$weight, data_small$height)
bmi_result
```




#6.Create a new variable to indicate which individuals are overweight, i.e. those with a BMI over 25.

```{r}
data_small_1['bmi_result_f']=bmi_result
data_small_1 %>% filter(bmi_result_f>25)

```

```{r}

data_small_1<- data_small_1 %>% mutate(overweight = if_else(bmi_result_f>25, 1, 0 ) )
data_small_1


```

##7.Some variables are treated as integers when they should be categorical variables. Change them into factors. 


```{r}
data_small$dob <- as.factor(data_small$dob)
data_small
```



8.Change levels of the gender variable, replacing 1 and 2 with female and male.


```{r}
data_small_1$gender <- cut(data_small_1$gender, 2, labels=c('female','male'))
data_small_1
```



##9.Group your data by gender and gluc. For each group, compute the mean and standard deviation of weight, then plot out a bar plot of computed means with error bars and use a facet to separate the data for the two genders.


```{r}

data_small %>% group_by(gender,gluc)%>%summarise(avg_weight = mean(weight),sd_weight = sd(weight))
```


```{r}

data_small %>% group_by(gender,gluc)%>%summarise(avg_weight = mean(weight),sd_weight = sd(weight))%>%  ggplot(aes(x=gender,y=avg_weight)) + geom_bar(stat="summary" ,col="blue2",fill="yellow",width=1)+stat_summary(fun = mean)+stat_summary(geom="errorbar")
```




#Part III – Plotting (5 points)

##10.Plot a jitter plot of the weight of individuals for both categories of active.


```{r}
data_small %>% ggplot(aes(x = active, y = weight))+ geom_jitter(aes(col =active),show.legend = TRUE)
```


##11.Plot out a grouped barplot of the number of individuals in the 3 categories of gluc, colored by active.


```{r}
data_small%>% group_by(gluc) %>% summarise(count_m_f=(n=n()))
```

```{r}

p1<- data_small_1%>% ggplot(aes(x=gluc,fill=gender, col=gender))+
  geom_bar(position="dodge")
p1


```
```{r}
data_small
```


```{r}

data_small%>%ggplot(aes(x=gluc,y=active))+
  geom_bar(stat = "identity")
```



12.Plot out a boxplot of age for both categories of alco.


```{r}
data_small_1
```


```{r}
library(patchwork)
library(gridExtra)
library(ggplot2)
library("dplyr")
```



```{r}
data_small_1$alco <- as.factor(data_small_1$alco)
data_small_1
# data_small
```

```{r}

data_small_1 %>% ggplot(aes(x=alco,y=age))+geom_boxplot(aes(col=alco))+
  ggtitle("boxplot of age for all categories of alco")

```

#13.Plot out a violin plot of the BMI for both males and females.


```{r}
data_small_1
```


```{r}
theme_set(theme_bw(base_size=12))
data_small_1 %>% ggplot(aes(y = bmi_result_f, x = gender),) + geom_violin(col="blue2",fill="yellow")
```





#Part IV – Selecting columns (3 point)

##14.Select the columns 1 and 4.

```{r}
data_small[c(1,4)]
```

15.Select all the columns ending with the letter “e”.

```{r}
data_small %>% select(ends_with('e'))

```
16.Select all the columns except smoke.

```{r}
data_small %>% select(-smoke)

```


#Part V – Filtering rows (4 point) 

##17.Filter the rows for males aged 64 and over.

```{r}
library(dplyr) 
```



```{r}
data_small_1 %>% filter(age>65 & gender=='male' )
```

##18.Filter the rows for alcoholics or smokers.


```{r}
data_small_1 %>% filter(alco==1 | smoke==1)
```

19.How many people smoke, drink and are active?

```{r}
nrow(data_small_1%>% filter(smoke ==1 & alco==1 & active==1))

```




20.How many patients are overweight? Show them by gender in a table.


```{r}
d1<-data_small_1[c("gender","overweight")]
d1%>% group_by(gender) %>% summarise(count_m_f=(n=n()))



```








