---
title: "Prabal Ghosh Workshop-3_R Notebook"
output:
  html_document:
    df_print: paged
---
#Workshop 3 : dplyr & ggplot
===
##Exercise 1: Filtering and selecting with dplyr

Read in the genomes.csv file into a variable called genomes.
 
Perform the following operations:
 
Make a list of the 5 organisms with the largest genomes 
 
Of the organisms which have more than 40 chromosomes which one comes first, alphabetically? 
 
Do any organisms containing a plasmid, also have more than one chromosome?
 
Make a version of the data containing only the columns from Chromosomes onwards.
 
Move the Size column to the front of the tibble.
 
Select just the columns which start with "O”.
 
How many different groups are there?  
===


```{r}
library(dplyr)
library(datasets)
```

```{r}
library(ggplot2)
```

```{r}
genom_data<- read.csv("C:\\Users\\praba\\Desktop\\uca1\\M1\\R programming\\class-6 workshop-3\\genomes.csv")
head(genom_data,10)
```
1.Make a list of the 5 organisms with the largest genomes 

```{r}
genom_data_1<- arrange(genom_data,desc(Size))
head(genom_data_1,5)
```
2.Of the organisms which have more than 40 chromosomes which one comes first, alphabetically? 


```{r}
data_1<-filter(genom_data, Chromosomes >= 40)
head(arrange(data_1,Organism),1)

```
3.Do any organisms containing a plasmid, also have more than one chromosome?

```{r}
genom_data%>% filter(Plasmids != 0 & Chromosomes>1)

```
4.Make a version of the data containing only the columns from Chromosomes onwards.

```{r}
# names(genom_data)
genom_data %>% select(Chromosomes:Assemblies)
```

```{r}
genom_data %>% select(4:7)
```
5.Move the Size column to the front of the tibble.

```{r}

genom_data_Sorted = genom_data %>% select("Size", everything()) 
genom_data_Sorted
```
6.Select just the columns which start with "O”.

```{r}
genom_data %>% select(starts_with('O'))
```
7.How many different groups are there? 

```{r}
genom_data %>% distinct(Groups) %>% nrow()
                                            
```


#Exercise 2: More clever filtering

Read in the “cancer_stats.csv” file into a variable called cancer then answer the following questions.
 
For which Digestive System cancer types are there more female cases than male cases?
 
Which cancer types (Class and Site) have no data (NA) for males? 
 
Which cancer types (Class and Site) have no data (NA) for females?
 
Which cancer type has the best survival rate for males? 
 
Which cancer Sites have “acute” in their names?
 
Out of tongue, kidney, breast and pancreas, which is classed as a soft tissue cancer?
===
Read in the “cancer_stats.csv” file into a variable called cancer then answer the following questions.

```{r}
cancer_data_1<- read.csv("C:\\Users\\praba\\Desktop\\uca1\\M1\\R programming\\class-6 workshop-3\\cancer_stats.csv")
head(cancer_data_1,15)
```
2.For which Digestive System cancer types are there more female cases than male cases?

```{r}

cancer_data_1 %>% filter(Class=="Digestive System") %>% filter(Female.Cases>Male.Cases) %>% select(Class, Site,Male.Cases,Female.Cases)
```
3.Which cancer types (Class and Site) have no data (NA) for males? 

```{r}
# cancer_data_1 %>% filter(Male.Cases == NA) %>% select(Class, Site)
cancer_data_1 %>% filter(is.na(Male.Cases)) 

```

4.Which cancer types (Class and Site) have no data (NA) for females?

```{r}
# cancer_data_1 %>% filter(Female.Cases == NA) %>% select(Class, Site)
cancer_data_1 %>% filter(is.na(Female.Cases)) 



```
5.Which cancer type has the best survival rate for males? 

```{r}
library(tidyr)

cancer_data_1 %>% drop_na(Male.Cases) %>% mutate(male_raio=Male.Deaths/Male.Cases) %>% arrange(male_raio)%>% slice(1)
```

6.Which cancer Sites have “acute” in their names?

```{r}
cancer_data_1 %>% filter(grepl('acute', Site))
library(stringr)
cancer_data_1 %>% filter(str_detect(Site,"acute"))

```


7.Out of tongue, kidney, breast and pancreas, which is classed as a soft tissue cancer?


```{r}
# vec1=c('tongue', 'kidney', 'breast','pancreas')
cancer_data_1 %>% filter(Class=='soft tissue') %>% filter(Site== 'tongue'| Site== 'kidney'| Site=='breast'|Site== 'pancreas' )
```
```{r}
vec1=c('tongue', 'kidney', 'breast','pancreas')
cancer_data_1 %>% filter(Class=='soft tissue') %>% filter(tolower(Site) %in% vec1 )
```



#Exercise 3: dplyr and ggplot

Load the starwars dataset.

Which variable (column) has the highest number of missing values?

How many humans contains the starwars dataset? show them by gender.

From which homeworld do the most individuals (rows) come from? 

Create a barplot of the gender distribution of the starwars Universe, set the title : "Gender distribution of the sw Universe". Make the colors of the columns depend on the gender, modify the colour using the command : scale_fill_manual.

Draw the densities for the height variable of feminines and masculines only.

Draw a segmented barplot for the variable 'sex'. The colors depend on the hair colours. Show the proportions [0,1] on the y-axis.
===

```{r}
data(starwars)
starwars
```
2.Which variable (column) has the highest number of missing values?

```{r}
a<-sort( colSums( sapply(starwars, is.na) ) )
a

```
2.How many humans contains the starwars dataset? show them by gender.


```{r}
# starwars %>% drop_na(gender,species) %>%filter(species == 'Human' & gender=='feminine')%>%summarize(n=n())
starwars%>%drop_na(gender,species) %>% count( (species == 'Human' & gender=='masculine'))
```
```{r}
starwars %>% drop_na(gender,species)%>% group_by(gender)  %>% count((species == 'Human' & gender=='masculine'))
```

```{r}
starwars %>% filter( species == 'Human' ) %>% group_by(gender) %>% summarise(count_m_f=(n=n()))
```

3.From which homeworld do the most individuals (rows) come from? 

```{r}
# starwars
starwars %>% group_by(homeworld) %>% summarise(count_homeworld=(n=n())) %>% arrange(desc(count_homeworld))

```
4.Create a barplot of the gender distribution of the starwars Universe, set the title : "Gender distribution of the sw Universe". Make the colors of the columns depend on the gender, modify the colour using the command : scale_fill_manual.


```{r}
starwars %>% drop_na(gender) %>% group_by(gender) %>% summarise(count_m_f=(n=n()))%>%ggplot(aes(x=gender,y=count_m_f),)+
  geom_bar(aes(fill=gender),stat = "identity", color = "brown") +
  labs(
    title = "Gender distribution of the sw Universe",
    xlab = "Gender",
    ylab= "Count"
  )+theme_light()
```

5.Draw the densities for the height variable of feminines and masculines only.


```{r}
starwars %>% drop_na(gender) %>% ggplot(aes(x=height,fill=gender))+geom_density(aes(col=gender))+theme_light()
```

6.Draw a segmented barplot for the variable 'sex'. The colors depend on the hair colours. Show the proportions [0,1] on the y-axis.

```{r}
# starwars

starwars %>% ggplot(aes(x=sex))+
  geom_bar(aes(fill=hair_color),position  = "fill", stat="count",col="Brown") +
  labs(
    title = "Bar Plot",
  )
```
```{r}
# starwars

starwars %>% drop_na(hair_color,sex) %>% ggplot(aes(x=sex))+
  geom_bar(aes(fill=hair_color),position  = "fill", stat="count",col="Brown") +
  labs(
    title = "Bar Plot",
  )
```

#Exercise 4: festival

Read in the festival.csv file into a variable called festival.

1.Restructure and clean up the data
	a- In which tidy format, the data are ? 
	b- Change this format in order to have observation per row and remove NA values (The 2 new columns must be day and cleanliness)

2. Draw a stripchart (geom_jitter) of the cleanliness values for males and females separately.
	a- Split the plot based on the day of the festival to see the effect this had on the data
	b- Personalize the point colors for males and females
	c- Add a line to show the mean

3. Add a new column called attendance to the data to say how many days people attended the festival. 

4. Now redraw the plot but faceting by both attendance and day


```{r}
festival_data<- read.csv("C:\\Users\\praba\\Desktop\\uca1\\M1\\R programming\\class-6 workshop-3\\festival.csv")
festival_data
```
##1.Restructure and clean up the data
	a- In which tidy format, the data are ? 
	b- Change this format in order to have observation per row and remove NA values (The 2 new columns must be day and cleanliness)
	
```{r}
library(tidyr)
```

```{r}
festival_data_tidy <- festival_data %>% pivot_longer(cols=day1:day3, names_to="day", values_to="cleanliness") %>% drop_na(cleanliness)
# dim(festival_data_tidy)
festival_data_tidy
```
```{r}
# dim(df_titanic_drop)
festival_data_drop <- festival_data %>% pivot_longer(cols=day1:day3, names_to="day", values_to="cleanliness") %>% na.omit()
# dim(festival_data_drop)
festival_data_drop
```


2. Draw a stripchart (geom_jitter) of the cleanliness values for males and females separately.
	a- Split the plot based on the day of the festival to see the effect this had on the data
	b- Personalize the point colors for males and females
	c- Add a line to show the mean
	
```{r}
festival_data_tidy<-festival_data %>% pivot_longer(cols=day1:day3, names_to="day", values_to="cleanliness") %>% drop_na(cleanliness)

festival_data_tidy %>%ggplot()+geom_jitter(aes(x=gender, y= cleanliness,col=gender))+
  geom_hline(yintercept = mean(festival_data_tidy$cleanliness, na.rm=TRUE), color='blue', lty='dashed', lwd=1)+
  facet_wrap(~ day, scales = "free")+
  labs(
    title= "stripchart of the cleanliness values for males and females",
  
  )+
  theme(legend.position="right")

```
	
	
###3. Add a new column called attendance to the data to say how many days people attended the festival. 

```{r}
festival_data_clean<- festival_data %>% pivot_longer(cols=day1:day3, names_to="day", values_to="cleanliness") %>% drop_na(cleanliness)
festival_data_clean_data<-festival_data_clean %>% group_by(person) %>% summarise(attendance=(n=n()))
festival_data_clean_data

```




###4. Now redraw the plot but faceting by both attendance and day

```{r}

festival_data_clean_data_join <- left_join(festival_data, festival_data_clean_data, by ='person')
festival_data_clean_data_join
```
```{r}
festival_data_clean_data_join <- left_join(festival_data, festival_data_clean_data, by ='person')

festival_data_clean_data_join %>%ggplot()+geom_jitter(aes(x=gender, y= attendance,col=gender))+
  labs(
    title= "the attendance values for males and females",
  
  )+
  theme(legend.position="right")
```
```{r}
festival_data_clean<- festival_data %>% pivot_longer(cols=day1:day3, names_to="day", values_to="cleanliness") %>% drop_na(cleanliness)
festival_data_clean_data_day<-festival_data_clean %>% group_by(day) %>% summarise(attendance=(n=n()))
festival_data_clean_data_day

```

```{r}

festival_data_clean_data_day %>%ggplot()+geom_bar(aes(x=day, y= attendance,fill=day),col='Brown',stat='identity')+
  labs(
    title= "bar plot of the attendance based on days",
  
  )+
  theme(legend.position="right")
```